Fast Neural Machine Translation Implementation

### 2018-05-18 wnmt.2018/wnmt.tex
Fix some grammar issues; add a reference to Tildes WMT17 paper

This paper describes the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante. We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation. We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam seach algorithm.

@@

We take an opposing approach by eschewing model selection and parameter tuning in favour of efficient implementation. We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models. As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a mean to explore the upper bound of the possible speed for a particular model and hardware. That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations. %The consequence of this focused approach can be seen in the results of the shared task ???.

We submitted two systems to this year's shared task for the efficient translation on GPU. Our first submission was tailored to be as fast as possible while being above the baseline BLEU score. Our second submission trades some of the speed of first submission to return better quality translations.

@@

The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that needs to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.

@@

We focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}. For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.

@@

Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found during the looping:
\begin{eqnarray*}
sum & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_b}} \\
    & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_a + \Delta}}  \\
    & = & e^{x_t - max_b} + e^{\Delta} \times \sum_{i=0...t-1}{e^{x_i - max_a}} 
\end{eqnarray*}
where $max_a$ is the previous
%!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
maximum value, $max_b$ is the now higher maximum value, ie. $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.

@@

In practise, however, efficiently using half-precision value requires a comprehensive redvelopment of the GPU code. We therefore make do with using the GPU's Tensor Core\footnote{https://devblogs.nvidia.com/programming-tensor-cores-cuda-9/} fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values and output a 32-bit float point product of the inputs.

@@

Both of our submitted systems use the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder. We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.

@@

Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models. 
As with our first submission, Amun is used as inference engine.
We trained 2 systems with differing vocabulary sizes and varied the beam sizes, and chose the configuration that produced the best results for translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.

>>>>

This paper describes the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante. We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation. We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam search algorithm.

@@

We take an opposing approach by eschewing model selection and parameter tuning in favour of efficient implementation. We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models. As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and it is useful as a means to explore the upper bound of the possible speed for a particular model and hardware. That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations. %The consequence of this focused approach can be seen in the results of the shared task ???.

We submitted two systems to this year's shared task for the efficient translation on GPU. Our first submission was tailored to be as fast as possible while being above the baseline BLEU score. Our second submission trades some of the speed of the first submission to return better quality translations.

@@

The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that need to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.

@@

We focus on the last three steps; their outline is shown in Algorithm~\ref{algo:original-softmax-beamsearch}. For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.

@@

Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found during the looping:
\begin{eqnarray*}
sum & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_b}} \\
    & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_a + \Delta}}  \\
    & = & e^{x_t - max_b} + e^{\Delta} \times \sum_{i=0...t-1}{e^{x_i - max_a}} 
\end{eqnarray*}
where $max_a$ is the previous
%!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
maximum value, $max_b$ is the now higher maximum value, i.e., $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.

@@

In practise, however, efficiently using half-precision value requires a comprehensive redevelopment of the GPU code. We therefore make do with using the GPU's Tensor Core\footnote{https://devblogs.nvidia.com/programming-tensor-cores-cuda-9/} fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values and output a 32-bit float point product of the inputs.

@@

Both of our submitted systems use a sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder. We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.

@@

Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models; multiplicative-LSTM's suitability for use in NMT models has been previously demonstrated by~\citet{pinnis2017}.
As with our first submission, Amun is used as inference engine.
We trained 2 systems with differing vocabulary sizes and varied the beam sizes, and chose the configuration that produced the best results for translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
